Improving Emotion Detection through Real-Time Translation of Text to ML Models Trained in Different Languages

### 2023-07-30 paper/appendix/dataset.tex
Paper Updates

# Rename Columns
df_english_filtered = df_english_filtered.withColumnRenamed("sentiment", "emotion")
df_english_filtered = df_english_filtered.withColumnRenamed("content", "sentence")

# Make sure the column order to the same for both german and english csv files
df_english_filtered = df_english_filtered.select('sentence', 'emotion')

@@

# Group By for Details & Count
df_english_grouped = df_english_filtered.groupBy('emotion').count()

@@

shutil.move(file_path, './data/data_en.csv')
\end{minted}
\clearpage

@@

# We Need to "split" the "artice_emotion" column since the ETH team listed
# multiple emotions in one column
# In order to "explode" the column to it's distinct rows
df_german_exploded = df_german.select('*', explode('article_emotion').alias('emotion'))

# Rename Column
df_german_exploded = df_german_exploded.withColumnRenamed("title", "sentence")

@@

# Make sure the column order to the same for both german and english csv files
df_german_exploded = df_german_exploded.select('sentence', 'emotion')

@@

# Group By for Details & Count
df_german_grouped = df_german_exploded.groupBy('emotion').count()

@@

shutil.move(file_path, './data/data_de.csv')
\end{minted}

\clearpage

>>>>

# Rename Columns
df_english_filtered = df_english_filtered.withColumnRenamed("sentiment", "emotion_en")
df_english_filtered = df_english_filtered.withColumnRenamed("content", "sentence_en")

# Make sure the column order to the same for both german and english csv files
df_english_filtered = df_english_filtered.select('sentence_en', 'emotion_en')

@@

# Group By for Details & Count
df_english_grouped = df_english_filtered.groupBy('emotion_en').count()

@@

shutil.move(file_path, './data/data_en.csv')

Count (raw): 40000
English Source Columns: ['tweet_id', 'sentiment', 'content']

Count (filtered): 35692

Grouped & Count by "emotion"
+----------+-----+
|emotion_en|count|
+----------+-----+
|love      |3842 |
|hate      |1323 |
|neutral   |8638 |
|anger     |110  |
|happiness |5209 |
|surprise  |2187 |
|sadness   |5165 |
|worry     |8459 |
|enthusiasm|759  |
+----------+-----+
\end{minted}
\clearpage

@@

# We Need to "split" the "artice_emotion" column since the ETH team listed
# multiple emotions in one column
# In order to "explode" the column to it's distinct rows
df_german_exploded = df_german.select('*', explode('article_emotion').alias('emotion_de'))

# Rename Column
df_german_exploded = df_german_exploded.withColumnRenamed("title", "sentence_de")

@@

# Make sure the column order to the same for both german and english csv files
df_german_exploded = df_german_exploded.select('sentence_de', 'emotion_de')

@@

# Group By for Details & Count
df_german_grouped = df_german_exploded.groupBy('emotion_de').count()

@@

shutil.move(file_path, './data/data_de.csv')

Count (raw): 1970
German Source Columns: ['article_emotion', 'article_id', 'article_stance', 'paragraphs', 'snippet', 'source', 'title']

Count (filtered): 2568

Grouped & Count by "emotion"
+------------+-----+
|emotion_de  |count|
+------------+-----+
|Vertrauen   |316  |
|Freude      |140  |
|Ärger       |226  |
|Überraschung|369  |
|Traurigkeit |184  |
|Antizipation|774  |
|Unklar      |314  |
|Angst       |154  |
|Ekel        |29   |
|Keine       |62   |
+------------+-----+
\end{minted}
\clearpage

\subsection{Details on Extending \& Translation}
\label{appendix:dataset_translation}
Details on translation and extending the datasets.
\begin{minted}[fontsize=\small]{python}
# **************************************
# *** English <-> German Emotion Key ***
# **************************************

from pyspark.sql.functions import col, udf, split, explode, lit
from pyspark.sql.types import IntegerType, StringType, StructType, StructField, StringType

# Emotion Dictionary English <-> German
# This key/value setup was previsoulsy created for linking purposes
emotion_key = {
    "boredom"    : "---",
    "love"       : "Vertrauen",
    "relief"     : "---",
    "fun"        : "---",
    "hate"       : "Ekel",
    "neutral"    : "Unklar",
    "anger"      : "Ärger",
    "happiness"  : "Freude",
    "surprise"   : "Überraschung", 
    "sadness"    : "Traurigkeit", 
    "worry"      : "Angst",
    "enthusiasm" : "Antizipation", 
    "empty "     : "---",
    "---"        : "Keine",
}

# Create the schema for the DataFrame
schema = StructType([
    StructField("emotion_en", StringType(), nullable=False),
    StructField("emotion_de", StringType(), nullable=False)
])

# Convert the dictionary to a list of tuples
data = [(key, value) for key, value in emotion_key.items()]

# Create the PySaprk DataFrame
df_emotion_key = sc.createDataFrame(data, schema)

df_emotion_key.show()

# Extend the English and German onto the datasets
df_german_exploded = df_german_exploded.join(df_emotion_key, on="emotion_de", how="left")
df_english_filtered = df_english_filtered.join(df_emotion_key, on="emotion_en", how="left")

# New sentence column
df_german_exploded = df_german_exploded.withColumn("sentence_en", lit(None))
df_english_filtered = df_english_filtered.withColumn("sentence_de", lit(None))

df_german_exploded.show()
df_english_filtered.show()

# Using Deep Translator to proxy to Google Translate
from deep_translator import GoogleTranslator

# Define Tranlation Functions
def translate_en_to_de(row):
    translated = GoogleTranslator(source='en', target='de').translate(row["sentence_en"])
    row["sentence_de"] = translated

def translate_de_to_en(row):
    translated = GoogleTranslator(source='de', target='en').translate(row["sentence_en"])
    row["sentence_en"] = translated

df_german_exploded.foreach(translate_de_to_en)

+----------+------------+
|emotion_en|  emotion_de|
+----------+------------+
|   boredom|         ---|
|      love|   Vertrauen|
|    relief|         ---|
|       fun|         ---|
|      hate|        Ekel|
|   neutral|      Unklar|
|     anger|       Ärger|
| happiness|      Freude|
|  surprise|Überraschung|
|   sadness| Traurigkeit|
|     worry|       Angst|
|enthusiasm|Antizipation|
|    empty |         ---|
|       ---|       Keine|
+----------+------------+

+------------+--------------------+----------+----------+-----------+----------+
|  emotion_de|         sentence_de|emotion_en|emotion_en|sentence_en|emotion_en|
+------------+--------------------+----------+----------+-----------+----------+
|   Vertrauen|Adoptiert zu sein...|      love|      love|       null|      love|
|      Freude|Österreichs Verfa...| happiness| happiness|       null| happiness|
|      Freude|Visana gewinnt er...| happiness| happiness|       null| happiness|
\end{minted}
\clearpage

### 2023-07-30 paper/main.tex
Paper Updates

\section{Introduction}
\subsection{Background on Human Emotion Theoretical Framework}

\begin{wrapfigure}{r}{0.4\textwidth} %this figure will be at the right
    \centering
    \includegraphics[width=0.35\textwidth]{Plutchik-Wheel}
    \caption{Plutchik Wheel of Emotions}
    \label{fig:plutchik-wheel}
\end{wrapfigure}

The concept of the eight main emotions can vary depending on the theoretical framework or model being considered. One well-known model is the Plutchik's Wheel of Emotions figure \ref{fig:plutchik-wheel}, which proposes eight primary emotions and their contrasting pairs. Here are the eight primary emotions according to the Plutchik's Wheel of Emotions \cite{Tromp}:

\begin{itemize}
\item Joy (Fun): A feeling of happiness, contentment, or delight.
\item Sadness: A state of unhappiness, sorrow, or grief.
\item Anger: An intense emotional response associated with feelings of displeasure, frustration, or hostility.
\item Worry: An emotional response triggered by perceived threats or danger, often associated with anxiety or terror.
\item Surprise: A sudden and unexpected reaction to something unusual or startling.
\item Disgust (Hate): A strong aversion or revulsion towards something unpleasant, offensive, or repulsive.
\item Trust: A sense of reliance, confidence, or faith in someone or something.
\item Anticipation: A feeling of excited expectation or eagerness towards future events.
\end{itemize}
It's important to note that emotions are complex and multifaceted, and different models may propose alternative categorizations or variations in the number of primary emotions. However for the scope of this research project the most common 8 emotions are used as described above. Both the English and German datasets being used in this research matched to the 8 listed above. Details on the linkage of the two datasets is described in the next section with an emphasis on the Table \ref{table:dataset_labels} detailed linkage.

@@

By accurately identifying and understanding emotions from text data, ML models can assist in improving user experiences, decision-making processes, and overall human-machine interactions in a positive manner\cite{Colonnello, mascarell-etal-2021-stance}. With the goal in mind of improving ML models this research paper came to be.

@@

\subsection{ED challenges, including lack of comprehensive labeled datasets and language fragmentation}
ED is growing field, however unlike Sentiment Analysis (SA) the availability of large datasets for training purposes of ML models is much smaller. The primary reason of the lack of data stems from the emotion detection requiring primarily supervised learning dataset, which require time and effort to procure, and mostly created by humans. To compound on the issue of the lack of data; the many datasets that are available are in multiple languages.

@@

This paper is significant because an entire industry of automated purported emotion reading (Text, Image, Video, and Sound) technologies is quickly emerging \cite{ACLU-ED-Data, ACLU-THE-DAWN-OF-ROBOT-SURVEILLANCE}. The market for emotion recognition software is forecast to reach at least \$3.8 billion by 2025 \cite{ACLU-ED-Data, ACLU-THE-DAWN-OF-ROBOT-SURVEILLANCE}. And since ED is already being incorporated into products for purposes such as marketing, robotics, driver safety, and a multiple of others, it is fitting to work on research to solve issues.

With the before mentioned issues regarding the quantity/lack and fragmentation of the ED datasets this research was motivated by finding a way to combine different language datasets into a single set for training processes to improve the predictability of ED. The hope is that by use of translation that the predictability of ED can be improved by use of real- real-time language conversion and processing.

@@

\begin{wrapfigure}{R}{0.4\textwidth} %this figure will be at the right
    \centering
    \includegraphics[width=0.35\textwidth]{Dataset-Extension}
    \caption{Example of English dataset extension by translating from German.}
    \label{fig:dataset-extension}
\end{wrapfigure}

@@

@@

\subsection{Data procurement process, including obtaining ED datasets in English and German}
The data procurement was relatively straight forward and in essence entailed using Google search terms for the Emotion Detection in English and German. Some of the main German data was obtained from the dataset built by ETH's Emotion and Stance Detection for German Text \cite{mascarell-etal-2021-stance} which also hosts a good website with their finds \footnote{https://mtc.ethz.ch/research/natural-language-processing/emotion-stance.html}. The English dataset was downloaded from Kaggle \footnote{https://www.kaggle.com/datasets/pashupatigupta/emotion-detection-from-text} based on Tweets collected in 2021. The English dataset contains over 38,000 tweets and their corresponding emotion labeled. Similarly the German contains over 12,000 sentences and their respective emotions associated.

@@

\subsection{Development of a translation application for converting different languages into English and vice versa}
Translation of the emotion datasets was performed with Google Translator based on the Medium article\cite{Nidhaloff_how_to_translate_text_with_python} in which the author describes option fro simple API translations. As an example the following is a brief introduction to Google Translator's API:

>>>>

\section{Introduction}

Emotion detection has gained significant interest in recent years as a means to understand human behavior and communication better. By analyzing text or speech, these detection processes can accurately identify and classify emotions expressed by individuals or groups. By employing a variety of ED processes, including natural language processing and machine learning, we can detect emotions with high precision. These ED systems find applications in diverse fields, such as sentiment analysis, customer feedback analysis, and mental health monitoring, contributing to a deeper understanding of human emotions and their impact on various aspects of life.

\subsection{Background on Human Emotion Theoretical Framework}

The concept of the eight main emotions can vary depending on the theoretical framework or model being considered. One well-known model is the Plutchik's Wheel of Emotions Figure \ref{fig:plutchik-wheel}, which proposes eight primary emotions and their contrasting pairs. Below are the eight primary emotions according to the Plutchik's Wheel of Emotions \cite{Tromp}, a detailed reference of emotions used in this research can be found in Table \ref{table:dataset_labels} on page~\pageref{table:dataset_labels}.

\begin{itemize}
\item Joy (Fun): A feeling of happiness, contentment, or delight.
\item Sadness: A state of unhappiness, sorrow, or grief.
\item Anger: An intense emotional response associated with feelings of displeasure, frustration, or hostility.
\item Worry: An emotional response triggered by perceived threats or danger, often associated with anxiety or terror.
\item Surprise: A sudden and unexpected reaction to something unusual or startling.
\item Disgust (Hate): A strong aversion or revulsion towards something unpleasant, offensive, or repulsive.
\item Trust: A sense of reliance, confidence, or faith in someone or something.
\item Anticipation: A feeling of excited expectation or eagerness towards future events.
\end{itemize}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{Plutchik-Wheel}
    \caption{Plutchik Wheel of Emotions}
    \label{fig:plutchik-wheel}
\end{figure}

It's important to note that emotions are complex and multifaceted, and different models may propose alternative categorizations or variations in the number of primary emotions. However for the scope of this research project the most common 8 emotions are used as described above. Both the English \cite{english-dataset-twitter} and German \cite{german-dataset-cheese} datasets being used in this research matched to the 8 listed above. Details on the linkage of the two datasets is described in Section \ref{sec:data-procurement-process} with an emphasis on the Table \ref{table:dataset_labels}'s detailed linkage.

@@

By accurately identifying and understanding emotions from text data, ML models can assist in improving user experiences, decision-making processes, and overall human-machine interactions in a positive manner\cite{Colonnello, mascarell-etal-2021-stance}. As an example, the primary goals of chatbots are entertainment, social contact, and novelty interaction, with a strong emphasis on productivity. In chatbots, there has been a tremendous requirement to simulate human-like characteristics and behavior.With the goal in mind of improving ML models this research paper came to be \cite{emotion-detection-literature-review}. These chatbots also improve customer engagement by offering friendliness, comfortness, flexibility, and efficient assistance. Chatbots should provide customers with more engaging responses, directly addressing their issues. Users perceive chatbots as companions rather than simple assistants. The majority of user requests are emotional rather than informative. Chatbots gained the capacity to respond emotionally to customers due to machine learning and sentiment analysis evolution\cite{chatbot-cognitive-awareness}.

@@

\subsection{ED challenges, including lack of comprehensive labeled datasets and language fragmentation}
ED is growing field, however unlike Sentiment Analysis (SA) the availability of large datasets for training purposes of ML models is much smaller \cite{ACLU-ED-Data}. The primary reason of the lack of data stems from the emotion detection requiring primarily supervised learning dataset, which require time and effort to procure, and mostly created by humans. To compound on the issue of the lack of data; the many datasets that are available are in multiple languages.

@@

This paper is significant because an entire industry of automated emotion reading (Text, Image, Video, and Sound) technologies is quickly emerging \cite{ACLU-ED-Data, ACLU-THE-DAWN-OF-ROBOT-SURVEILLANCE}. The market for emotion recognition software is forecast to reach at least \$3.8 billion by 2025 \cite{ACLU-ED-Data, ACLU-THE-DAWN-OF-ROBOT-SURVEILLANCE}. And since ED is already being incorporated into products for purposes such as marketing, robotics, driver safety, customer service and a multiple of others, it is fitting to work on research to solve issues with prediction rates and lack of teaching / learning data.

With the before mentioned issues regarding the quantity (lack of) and fragmentation of the ED datasets this research was motivated by finding a way to combine different language datasets into a single set for training processes to improve the predictability of ED. The hope is that by use of translation that the predictability of ED can be improved by use of real-time language conversion and processing.

@@

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.35\textwidth]{Dataset-Extension}
    \caption{Example of English dataset extension by translating from German.}
    \label{fig:dataset-extension}
\end{figure}

@@

\subsection{Data procurement process, including obtaining ED datasets in English and German}
\label{sec:data-procurement-process}
The data procurement was relatively straight forward and in essence entailed using Google search terms for the Emotion Detection in English and German. Some of the main German data was obtained from the dataset built by ETH's Emotion and Stance Detection for German Text \cite{mascarell-etal-2021-stance} which also hosts a good website with their finds \footnote{https://mtc.ethz.ch/research/natural-language-processing/emotion-stance.html}. The English dataset was downloaded from Kaggle \footnote{https://www.kaggle.com/datasets/pashupatigupta/emotion-detection-from-text} based on Tweets collected in 2021. The English dataset contains over 38,000 tweets and their corresponding emotion labeled. Similarly the German contains over 12,000 sentences and their respective emotions associated.

@@

\subsection{Development of a translation application for converting different languages into English and vice versa}
Translation of the emotion datasets was performed with Google Translator based on the Medium article\cite{Nidhaloff_how_to_translate_text_with_python} in which the author describes option for simple API translations. As an example the following is a brief introduction to Google Translator's API:
